# Problem Set 3

## Single-Agent Dynamic Optimization

### Question 1
This equation represents the maximum expected discounted lifetime utility for an agent observed in state $x$ for a given choice of whether to replace the bus engine $i \in \{0,1\}$:

$$EV^1(x,i) = \int_y \log \left\{\sum_{j \in C(y)} \exp[u(y, j;\theta) + \beta EV^0(y,j)]\right\}p(dy|x,i)$$

Because this is a contraction mapping, $EV_{\theta}$ is the unique solution to this equation under $EV^1 = EV^0$.

The second equation is the multinomial choice probability of option $i$ given observed state $x_t$ and parameter vector $\theta$:

$$P(i_t|x_t;\theta) = \frac{\exp(u(x_t, i_t;\theta) + \beta EV(x_t,i_t;\theta))}{\sum_{i = 0,1}\exp(u(x_t,i;\theta) + \beta EV(x_t,i;\theta))}$$

```{r results='hide'}
rm(list = ls())

# Question 1(ii) ----
# * Initialize parameters
RC <- 9.7558
c <- 2.6275
p <- .3489
q <- .6394
beta <- .9999

mileage_bins <- seq(from = 0, to = 89, by = 1)

# Initialize the transition matrix
transition_matrix <- matrix(0, nrow = 90, ncol = 90)

# Populate the transition matrix with probabilities
for (i in 1:90) {
    if (i == 89) {
        transition_matrix[i, i] <- p
        transition_matrix[i, i + 1] <- 1 - q
    } else if (i == 90) {
        transition_matrix[i, i] <- 1
    } else {
        transition_matrix[i, i] <- p
        transition_matrix[i, i + 1] <- q
        transition_matrix[i, i + 2] <- 1 - p - q
    }
}
```

```{r}
library(data.table)
library(ggplot2)

# Expected value function initialization
EV <- rep(1, 90)
c_vec <- 0:89 * c

# Bellman iteration to calculate expected value
for (i in 1:1000) {
    EV_new <- transition_matrix %*%
        (log(exp(-c_vec + beta * EV) + exp(-RC + beta * EV[1])))

    # Recentering step: subtract the minimum EV to prevent numerical issues
    EV_new <- EV_new - min(EV_new)

    if (max(abs(EV_new - EV)) < 1e-6) { # Convergence check
        print(paste0("Converged at iteration ", i))
        break
    }

    EV <- EV_new
}

# Plot the expected value function
dt <- data.table(mileage_bins, EV)

ggplot(dt, aes(x = mileage_bins, y = EV)) +
    geom_line() +
    labs(x = "Mileage", y = "Expected Value", title = "Expected Value Function") +
    theme_classic()
```

### Question 2
Replicating Rust (1987).

```{r}
library(here)
library(data.table)
library(R.matlab)

# Load data ----
dt <- as.data.table(readMat(here("Mortimer HW", "data", "ps3_data.mat")))
names(dt) <- c("bus_id", "group", "year", "month", "mileage", "mileage_diff")

# Estimate transition probabilities
dt[, mileage_int := floor(mileage / 5000)]
dt[, mileage_int_lag := shift(mileage_int, 1L, type = "lag"), by = bus_id]
dt[is.na(mileage_int_lag), mileage_int_lag := mileage_int]
dt[, mileage_diff_int := mileage_int - mileage_int_lag]

# Determine when engine is replaced
dt[, d := fifelse(mileage_diff_int < 0, 1, 0)]

p <- nrow(dt[mileage_diff_int == 0]) / nrow(dt[mileage_diff_int >= 0])
q <- nrow(dt[mileage_diff_int == 1]) / nrow(dt[mileage_diff_int >= 0])

# Bellman iteration to calculate expected value given x = 0, 1, ..., 89
ev_func <- function(theta) {
    RC <- theta[2]
    c_vec <- theta[1] * 0:89

    EV <- rep(0, 90)
    for (i in 1:1000) {
        # Compute the max to use in the Log-Sum-Exp trick
        max_val <- pmax(-c_vec + EV, -RC + EV[1])

        # Compute the stable log-sum-exp expression
        EV_new <- transition_matrix %*%
            (max_val + log(exp(-c_vec + EV - max_val) +
                exp(-RC + EV[1] - max_val)))

        if (any(is.na(EV_new))) {
            warning(paste0("NA values found at iteration ", i))
            print(theta)
            break # Exit the loop if NA values are found
        }

        # Recentering step: subtract the minimum EV to prevent numerical issues
        EV_new <- EV_new - min(EV_new)

        if (max(abs(EV_new - EV)) < 1e-6) { # Convergence check
            print(paste0("Converged at iteration ", i))
            return(EV_new)
        }

        EV <- EV_new
    }
}

# Define partial likelihood function
P_0 <- function(theta, d_data, x_data) {
    RC <- theta[2]
    c <- theta[1]

    EV <- ev_func(theta)

    denom <- (1 + exp(x_data * c - RC - beta * EV[x_data + 1] + beta * EV[1]))
    return(abs(d_data - 1 / denom))
}

L_p <- function(theta, d, x) {
    return(-sum(log(P_0(theta, d, x))))
}

theta <- optim(c(5, 5), L_p, d = dt$d,
    x = dt$mileage_int, method = "Nelder-Mead")
print(theta$par)
```

The contraction mapping has issues when searching over very negative values of $c$ and $RC$. This is because exponentiating large values overflows the memory and results in NaN values. While my replacement cost parameter is close-ish to Rust, I am unable to get a positive maintenance cost parameter.

