---
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

# Empirical IO: HW4
# Mo Van De Sompel, Makoto Tanaka, and Colin Williams
Estimate Brenehan and Reiss (1990, 1991) model of firm entry using airline data.

```{r, echo = FALSE}
rm(list = ls())

dir <- dirname(dirname(rstudioapi::getSourceEditorContext()$path))
setwd(dir)

pacman::p_load(data.table, haven, knitr)

dt <- as.data.table(read_dta("HW3/airline1997data.dta"))
```

## Aggregate data to market level

```{r, include = TRUE}
dt <- unique(dt[tkcarrier %in% c("AA", "CO", "DL", "NW",  "TW", "UA", "US"), 
                .(market, tkcarrier, marketdistanceticket)])
dt[, nFirms := .N, by = .(market)]

dt.market <- unique(dt[, .(nFirms, market, marketdistanceticket)])
```


## Estimate profit function paramters via maximum likelihood
Define the reduced form profit function $$\pi_{im}(N_m) = X_m\alpha + g(N_m) + \epsilon_{0m}$$

where $$g(N_m) = \sum_{j=1}^7\theta_j\mathbf{1}\{N_m = j\}$$

1. The $\alpha$ parameter captures how the profitability of serving a market varies with the flight distance. The $\theta$ parameters measure how the entry of additional firms affect per-firm profits.

2. We make several assumptions to estimate this model:
- All firms experience an identical market-level profitability shock, $\epsilon_{im} = \epsilon_{0m}$
- All firms have the same loss in profits due to flight distance, $\alpha_i = \alpha$
- All firms have an identical effect on their rivals' profits when they enter any market, $\theta_{ijm} = \theta_j$
- More entrants reduce firm profitability, $\theta_j > \theta_k$ for $j < k$.

These symmetry assumptions allow the econometrician to model the number, rather than the identity, of entering firms.

3. Model estimates:

```{r, echo = FALSE, warnings = FALSE}
Profits <- function(params, dist, nFirms) {
  alpha <- params[1]
  thetas <- params[2:8]
  
  return(dist*alpha + thetas[nFirms])
}

Prob <- function(params, dist, nFirms) {
  if (nFirms == 0) {
    return(pnorm(-Profits(params, dist, 1), log.p = TRUE))
  } else if (nFirms < 7) {
    return(log(pnorm(Profits(params, dist, nFirms)) - pnorm(Profits(params, dist, nFirms+1))))
  } else {
    return(pnorm(Profits(params, dist, nFirms), log.p = TRUE))
  }
}

Likelihood <- function(params, dist, nFirms) {
  dt.est[, f := Prob(params, marketdistanceticket, nFirms), by = seq_len(nrow(dt.est))]  
  return(-sum(dt.est$f))  
}

guess <- c(-.005, seq(-.1, -.8, -.1))
dt.est <- copy(dt.market)
optim <- optim(par = guess, Likelihood)
```

```{r, include = TRUE}
dt.optim <- data.table(Parameter = c("alpha", paste0("theta", 1:7)), Estimate = optim$par)
kable(dt.optim)
```


